\dnest~Manual

### 2016-01-13 doc/manual.tex
Working on the manual to make it correct for DNest4

\dnest~is a multi-threaded C++ implementation of Diffusive Nested Sampling
(DNS), a powerful Markov Chain Monte Carlo (MCMC) algorithm that is mainly
useful for solving Bayesian Inference problems.
\dnest~is free software,
released under the terms of the GNU General Public Licence, version 3.
For information about how the
algorithm works, please see the paper, freely available from the following
URL:\\

@@

If you find \dnest~useful, please feel free to cite the paper.
Throughout this manual I will assume that you have read the paper and have a
reasonable understanding of how the algorithm works, as well as some
understanding of Nested Sampling in general, as introduced by John Skilling.\\

@@

\subsection{Dependencies}
\dnest~depends on some other software that you will need to have installed on
your system. These should be quite easy to obtain. I highly
recommend obtaining these programs from your operating system's package manager,
rather than installing them from the source code. Note that some OSs split up
packages into binaries (the library files) and ``development'' packages,
containing header files (part of a library's source code).
In Ubuntu, Linux Mint, and Debian, the development packages have a -dev suffix.
In Fedora, the development packages usually have a -devel suffix (e.g. Fedora).
To compile \dnest, you will need both the libraries and the development packages
for each of the dependencies. Here's a list of dependencies:\\

\begin{itemize}
\item The GNU Scientific Library (GSL)
\item Python 2, along with the NumPy and matplotlib packages
\item The {\tt zlib} library (this is only required if wanting the ability to
output samples to a {\tt gzipped} file, otherwise it can be excluded)
\item The {\tt thread}, {\tt system} and {\tt iostreams} parts of the Boost C++ library.
It may be easier to just install all of Boost.
\item To compile \dnest, you need {\tt cmake}.\\
\end{itemize}

On Ubuntu, I can install these packages with the following commands:
\begin{verbatim}
sudo apt-get install libgsl0-dev
sudo apt-get install python-numpy python-matplotlib
sudo apt-get install zlib1g-dev
sudo apt-get install libboost-thread-dev libboost-system-dev libboost-iostreams-dev
sudo apt-get install cmake
\end{verbatim}

\dnest~uses GSL's random number generator. Python, NumPy and
matplotlib are used for the postprocessing scripts and for
plotting. Boost is used for parallelization (multithreading).
While random number generation and
multithreading is supported
in the recent C++11 standard, I have not ported \dnest~to C++11 and have no
intention of doing so in the near future.\\

It is also possible to compile \dnest~without Boost, however this is not
recommended, and you will not be able to use any of the multi-threaded options
or the {\tt gzip} option.\\

\subsection{Compiling}
Please follow the instructions in README.md to compile \dnest~using cmake.
This is the recommended option.\\

%Go into the \dnest~directory and type {\tt make}. Everything should compile.
To
verify that the compilation was successful,
go into the {\tt Examples/SpikeSlab} directory
and see if an executable file called {\tt main} has been created. If you
execute this file, you should see a lot of output about ``levels'' and
``particles'' (see Figure~\ref{fig:output} for example output).
You will need to manually terminate this process by using
Control+C, or whatever works in your operating system. If you saw this output,
\dnest~was compiled successfully.\\

>>>>

\dnest~is a multi-threaded C++ implementation of Diffusive Nested Sampling
(DNS), a powerful Markov Chain Monte Carlo (MCMC) algorithm that is mainly
useful for solving Bayesian Inference problems.
\dnest~is free software, released under the terms of the MIT licence.
For information about how the
algorithm works, please see the Diffusive Nested Sampling paper,
available from the following URL:\\

@@

If you find \dnest~useful in a research project,
please feel free to cite the paper.
Throughout this manual I will assume that you have read the paper and have a
reasonable understanding of how the algorithm works, as well as some
understanding of general Nested Sampling concepts.\\

@@

\subsection{Dependencies}
One of the issues with DNest3 which inspired a new version was that many users
found it annoying to compile on different operating systems because of the
dependencies. For that reason, \dnest~now has much less dependencies.
The C++ part of the code is written in pure C++11 and does not require any
additional libraries --- you just have to have a modern C++ compiler that
supports the C++11 standard.

The Python postprocessing scripts use the Python packages
{\tt numpy} and {\tt matplotlib}. They work with both Python 2 or 3.

\subsection{Compiling}
Go into the {\tt code} directory and run {\tt make}. Everything should compile.
To
verify that the compilation was successful,
go into the {\tt Examples/SpikeSlab} directory
and see if an executable file called {\tt main} has been created. If you
execute this file, you should see a lot of output about ``levels'' and
``particles'' (see Figure~\ref{fig:output} for example output).
You will need to manually terminate this process by using
Control+C, or whatever works in your operating system. If you saw this output,
\dnest~was compiled successfully.\\
